---
title: "Lab 6"
author: "Van Gonzales"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann- Whitney U (non-parametric alternative to an unpaired 2-sample t-test)

H0: Equal ranks (or medians)
HA: Ranks are not equal (or medians)

### Create samples

`set.seed()` to create "pseudorandom, but reproducible samples"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributions using `hist()`

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2)

# If these data were paired, I can just add the argument 'paired = "TRUE"' within the t.test()
```

This p-value means that if our samples are drawn from pops with the same mean (if the null is true), there is `r round(ttest$p.value, 2)` probability of finding sample means *at least as different* as mine by random chance.

[A more comprehensive explanation of difference between groups...actually means data spread, effect size]. there is no sig diff (t(df) = t-statistic vale, p = ).

### Mann-Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that if our samples are drawn from populations with the same median (if the null is true), there is `r round(mwu$p.value, 2)` probability of finding sample medians *at least as different* as mine by random chance.

No significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p = `r round(mwu$p.value, 2)`).



## Simple Linear Regression

Two *continuous quantitative* variables that we are trying to understand the relationship between and we think that a linear relationship makes sense.

Use `penguiins` data from {palmerpenguins}, exploring flipper length (x) versus body mass(y).

```{r}
# Use view(penguins) to look at it
#Makes an exploratory scatterplot of penguiin flipper length (X-AXIS) versus body mass (y-axis)

ggplot(data = penguins, aes(x = flipper_length_mm, y= body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")
```


### Find the model estimate (coefficients)

Use the `lm()` function for the linear model.

~ <- that is a tilde not a minus sign -- body mass (g) = 49.69(flipper length) - 5780.83-- 49.69g/mm -> 49.69 gram increase in body mass per 1mm increase of flipper length-- -5780.8 means that is the expected average mass if the flipper length is zero
```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```


Model coefficients: `r penguin_lm$coefficients`
Flipper length coefficient: `r penguin_lm$coefficients[2]`

### Use `broom::tidy()` to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy

penguin_lm_tidy$estimate[2]
```
### Use broom::glance() to get output of model fit and statistical significance

```{r}
penguin_lm_glance <- glance(penguin_lm)
penguin_lm_glance
```
### Use equatiomatic::extract_eq() to return model in LaTeX

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

### Use plot() function to automatically produce diagnostic plots

```{r}
plot(penguin_lm)
```


